Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-09 paper_jasa.tex
still lemma 3.1

\begin{eqnarray*}
P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt = 1 - E\left[ (1 - p_D)^{m-1}\right],
\end{eqnarray*}

}
%
%
%
%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
%\begin{eqnarray*}
%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
%\end{eqnarray*}
%which, in particular, is independent of $p_{0,i}$ for all $i$.
%
%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all observers  have the same ability in identifying the data plot.
%
%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
%
%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
%\[
%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
%\]
\end{proof}

@@

Now assume that the same lineup plot is evaluated by $K$ independent observers. W.l.o.g.\green{Can we get rid of cliches like WLOG?} let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.  Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.

For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
we get
\[
N_m\sim \left \{ 
\begin{array}{ll}
B_{K, (1-E[p_D])^{m-1}} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]

>>>>

\begin{eqnarray*}
P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt = 1 - E\left[ (1 - p_D)^{m-1}\right] = \\
&=& 1 - \sum_{k=0}^{m-1} {m-1 \choose k} (-E[p_D])^{k} = (m-1)E[p_D] + O(E[p_D^2]),
\end{eqnarray*}

This can't be right - it's completely counter-intuitive: as $p_D$, the $p$-value is smaller, the probability that we pick the data plot should go up not down. 
Conversely, as $m$, the number of choices given, goes up the probability to pick the data plot should go down, not up. 
}
%
%
%
%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
%\begin{eqnarray*}
%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
%\end{eqnarray*}
%which, in particular, is independent of $p_{0,i}$ for all $i$.
%
%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all observers  have the same ability in identifying the data plot.
%
%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
%
%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
%\[
%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
%\]
\end{proof}

@@

Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
Let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.  Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.

For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
we get
\[
N_m\sim \left \{ 
\begin{array}{ll}
B_{K, 1-E[1 - p_D])^{m-1}} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]
